---
title: "In-Class Exercise 07: Calibrating Hedonic Pricing Model for Private Highrise Property with GWR Method"
author: "Pelle Knegjes"
format: html
editor: visual
date: "Okt 14 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

*LESSON NOTE*

To turn multi into single poly

TMSHAPE

sf_polygon = prov_sf

st(cast (polygon

mutate area = st_area(.))

prov_vleaned = sfpolygon

groupby province

filter area == max(area

ungroup()

select(-area)

select(province)

------------------------------------------------------------------------

```{r}
pacman::p_load(olsrr, corrplot, performance, ggpubr, sf, spdep, sfdep, GWmodel, tmap, tidyverse, glue, gtsummary, ggstatsplot)
```

```{r}
mpsz = st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_WEB_PL")
```

```{r}
mpsz_svy21 <- st_transform(mpsz, 3414)
```

```{r}
condo_resale = read_csv("data/aspatial/Condo_resale_2015.csv")
```

```{r}
glimpse(condo_resale)
```

```{r}
condo_resale.sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

```{r}
corrplot(cor(condo_resale[, 5:23]), diag = FALSE, order = "AOE",
         tl.pos = "td", tl.cex = 0.5, method = "number", type = "upper")

```

```{r}
#| fig-width: 12
#| fig-height: 10
ggcorrmat(condo_resale[, 5:23])
```

```{r}
condo.mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                   PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                   PROX_URA_GROWTH_AREA + PROX_MRT  + PROX_PARK + 
                   PROX_PRIMARY_SCH + PROX_SHOPPING_MALL    + PROX_BUS_STOP + 
                   NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD + LEASEHOLD_99YR,
                 data=condo_resale.sf)
ols_regress(condo.mlr)

```

EXCLUDE STATISTICLAYY INSIGNIFICANT IN RESUTLS\
Sign \<0.5

VIF between 0 and 5 no problem

above 5 meh

above 10 no go

```{r}
ols_vif_tol(condo.mlr)
```

### variable selection

```{r}
condo_fw_mlr = ols_step_forward_p(
  condo.mlr,
  p_val = 0.05,
  details = FALSE
)
```

```{r}
plot(condo_fw_mlr)
```

```{r}
ggcoefstats(condo.mlr,
            sort = "ascending")
```

```{r}
ols_plot_resid_fit(condo_fw_mlr$model)
```

```{r}
ols_plot_resid_hist(condo_fw_mlr$model)
```

```{r}
ols_test_normality(condo_fw_mlr$model)
```

```{r}
mlr_output = as.data.frame(condo_fw_mlr$model$residuals)%>%
  rename('FW_MLR_RES' = 'condo_fw_mlr$model$residuals')
```

```{r}
condo_resale_sf = cbind(condo_resale.sf,
                        mlr_output$FW_MLR_RES) %>%
  rename('MLR_RES' = 'mlr_output.FW_MLR_RES')
```

```{r}
tmap_mode("plot")
tmap_options(check.and.fix = TRUE)
tm_shape(mpsz)+
  tmap_options(check.and.fix = TRUE) +
  tm_polygons(alpha = 0.4) +
  tm_shape(condo_resale_sf) +
  tm_dots(col = "MLR_RES",
          alpha = 0.6,
          style = "quantile")

  
```

```{r}
tmap_mode("plot")
```

```{r}
condo_resale_sf = condo_resale_sf %>%
  mutate(nb = st_knn(geometry, k=6,
                     longlat = FALSE),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1)
```

```{r}
global_moran_perm(condo_resale_sf$MLR_RES,
                  condo_resale_sf$nb,
                  condo_resale_sf$wt,
                  alternative = "two.sided",
                  nsim = 99)
```

Null hypothesis that its random is rejected so there is clustering

```{r}
bw_fixed = bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE + PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE + PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + PROX_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
                  data=condo_resale_sf,
                  approach = "CV",
                  kernel = "gaussian",
                  adaptive = FALSE,
                  longlat = FALSE)

```

use the bandwith in the calibration

```{r}
gwr_fixed = gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE + PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE + PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + PROX_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD,
                  data=condo_resale_sf,
                  bw=bw_fixed,
                  kernel = "gaussian",
                  longlat = FALSE)
```

```{r}
gwr_fixed
```

```{r}
bw_adaptive <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM + AGE  + 
                        PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE    + 
                        PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                        PROX_PRIMARY_SCH + PROX_SHOPPING_MALL   + PROX_BUS_STOP + 
                        NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                      data=condo_resale_sf, 
                      approach="CV", 
                      kernel="gaussian", 
                      adaptive=TRUE, 
                      longlat=FALSE)
```

```{r}
gwr_adaptive <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM + AGE + 
                            PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE + 
                            PROX_URA_GROWTH_AREA + PROX_MRT + PROX_PARK + 
                            PROX_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_BUS_STOP + 
                            NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                          data=condo_resale_sf, 
                          bw=bw_adaptive, 
                          kernel = 'gaussian', 
                          adaptive=TRUE, 
                          longlat = FALSE)
```

```{r}
gwr_adaptive
```

```{r}
gwr_adaptive_output <- as.data.frame(
  gwr_adaptive$SDF) %>%
  select(-c(2:15))
```

```{r}
gwr_sf_adaptive <- cbind(condo_resale_sf,
                         gwr_adaptive_output)
```

```{r}
glimpse(gwr_sf_adaptive)
```

```{r}
summary(gwr_adaptive$SDF$yhat)
```

```{r}
tmap_mode("plot")
tmap_options(check.and.fix = TRUE)
tm_shape(mpsz)+
  tm_polygons(alpha = 0.1) +
tm_shape(gwr_sf_adaptive) +  
  tm_dots(col = "Local_R2",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))
```

```{r}
tmap_options(check.and.fix = TRUE)
tmap_mode("plot")
AREA_SQM_SE <- tm_shape(mpsz)+
  tm_polygons(alpha = 0.1) +
tm_shape(gwr_sf_adaptive) +  
  tm_dots(col = "AREA_SQM_SE",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

AREA_SQM_TV <- tm_shape(mpsz)+
  tm_polygons(alpha = 0.1) +
tm_shape(gwr_sf_adaptive) +  
  tm_dots(col = "AREA_SQM_TV",
          border.col = "gray60",
          border.lwd = 1) +
  tm_view(set.zoom.limits = c(11,14))

tmap_arrange(AREA_SQM_SE, AREA_SQM_TV, 
             asp=1, ncol=2,
             sync = TRUE)

```

```{r}
tmap_mode("plot")
```

```{r}
tm_shape(mpsz[mpsz$REGION_N=="CENTRAL REGION", ])+
  tm_polygons()+
tm_shape(gwr_sf_adaptive) + 
  tm_bubbles(col = "Local_R2",
           size = 0.15,
           border.col = "gray60",
           border.lwd = 1)
```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
