---
title: "Take Home Exercise 1"
author: "Pelle Knegjes"
format: html
editor: visual
date: "Aug 27 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
  warning: false
---

# **Take-home Exercise 1: Geospatial Analytics for Public Good**

## **Setting the Scene**

According to World Health Organisation (WHO), road traffic accidents cause the death of approximately 1.19 million people each year leave between 20 and 50 million people with non-fatal injuries. More than half of all road traffic deaths occur among vulnerable road users, such as pedestrians, cyclists and motorcyclists.

Road traffic injuries are the leading cause of death for children and young adults aged 5–29. Yet two thirds of road traffic fatalities occur among people of working age (18–59 years). Nine in 10 fatalities on the roads occur in low- and middle-income countries, even though these countries have around 60% of the world’s vehicles.

In addition to the human suffering caused by road traffic injuries, they also incur a heavy economic burden on victims and their families, both through treatment costs for the injured and through loss of productivity of those killed or disabled. More broadly, road traffic injuries have a serious impact on national economies, costing countries 3% of their annual gross domestic product.

Thailand’s roads are the deadliest in Southeast Asia and among the worst in the world, according to the World Health Organisation. About 20,000 people die in road accidents each year, or about 56 deaths a day (WHO).

Between 2014 and 2021, Thailand experienced a notable increase in accident frequencies. Specifically, 19% of all accidents in Thailand occurred on the national highways, which constituted the primary public thoroughfares connecting various regions, provinces, districts, and significant locations within a comprehensive network. Within the broader context of accidents across the country, there existed a considerable 66% likelihood of encountering accident-prone zones, often termed ‘black spots,’ distributed as follows: 66% on straight road segments, 13% at curves, 6% at median points of cross-shaped intersections, 5% at T-shaped intersections and Y-shaped intersections, 3% at cross-shaped intersections, 2% on bridges, and 2% on steep slopes, respectively.

## **Objectives**

By and large, road traffic accidents can be attributed by two major factors, namely: behavioural and environmental factors. Behavioural factors in driving are considered to be major causes of traffic accidents either in direct or indirect manner (Lewin, 1982). These factors can be further grouped into two as, driver behavior (also called driver/driving style) and driver performance, in other words, driver/driving skills (Elander, West, & French, 1993). Environmental factors, on the other hand, includes but not limited to weather condition such as poor visibility during heavy rain or foggy and road conditions such as sharp bend road, slippery slope road, and [blind spot](https://en.wikipedia.org/wiki/Vehicle_blind_spot).

Previous studies have demonstrated the significant potential of Spatial Point Patterns Analysis (SPPA) in exploring and identifying factors influencing road traffic accidents. However, these studies often focus solely on either behavioral or environmental factors, with limited consideration of temporal factors such as season, day of the week, or time of day.

In view of this, you are tasked to discover factors affecting road traffic accidents in the [Bangkok Metropolitan Region BMR](https://en.wikipedia.org/wiki/Bangkok_Metropolitan_Region) by employing both spatial spatio-temporal point patterns analysis methods.

The specific objectives of this take-home exercise are as follows:

-   To visualize the spatio-temporal dynamics of road traffic accidents in BMR using appropriate statistical graphics and geovisualization methods.

-   To conduct detailed spatial analysis of road traffic accidents using appropriate Network Spatial Point Patterns Analysis methods.

-   To conduct detailed spatio-temporal analysis of road traffic accidents using appropriate Temporal Network Spatial Point Patterns Analysis methods.

## **The Data**

For the purpose of this exercise, three basic data sets must be used, they are:

-   [Thailand Road Accident \[2019-2022\]](https://www.kaggle.com/datasets/thaweewatboy/thailand-road-accident-2019-2022) on Kaggle

-   [Thailand Roads (OpenStreetMap Export)](https://data.humdata.org/dataset/hotosm_tha_roads) on HDX.

-   [Thailand - Subnational Administrative Boundaries](https://data.humdata.org/dataset/cod-ab-tha?) on HDX.

# 2 Getting Started

## 2.1 Objectives

The specific objectives of this take-home exercise are as follows:

-   To visualize the spatio-temporal dynamics of road traffic accidents in BMR using appropriate statistical graphics and geovisualization methods.

-   To conduct detailed spatial analysis of road traffic accidents using appropriate Network Spatial Point Patterns Analysis methods.

-   To conduct detailed spatio-temporal analysis of road traffic accidents using appropriate Temporal Network Spatial Point Patterns Analysis methods.

## 2.2 Area of Study

Study Area is Bangkok Metropolitan Region

WGS 84 UTM zone 47N EPSG code of bangkok: 32647


## 2.3 Loading the tools needed in R

The code chunk below uses pacman to install and load the sf, raster, spatstat, tmap, tidyverse, sparr, and spNetwork tools

```{r}
pacman::p_load(sf, raster, spatstat, sparr, tmap, tidyverse, spNetwork)
```

Explanations for the imported library:

-   [sf](https://r-spatial.github.io/sf/) for handling geospatial data

-   [spatstat](empty)

-   [tmap](https://cran.r-project.org/web/packages/tmap/) for visualizing geospatial

-   [knitr](https://www.rdocumentation.org/packages/knitr/versions/1.45) for creating html tables

-   [Hmisc](https://www.rdocumentation.org/packages/Hmisc/versions/5.1-1) for summary statistics


-   [DT](https://rstudio.github.io/DT/) library to create interactive html tables

-   [**raster**](https://rspatial.org/raster/) for handling raster data in R,

-   [**spatstat**](https://cran.r-project.org/web/packages/spatstat/index.html) for performing Spatial Point Patterns Analysis such as kcross, Lcross, etc.,

-   [**sparr**](https://tilmandavies.github.io/sparr/index.html) provides functions to estimate fixed and adaptive kernel-smoothed spatial relative risk surfaces via the density-ratio method and perform subsequent inference. Fixed-bandwidth spatiotemporal density and relative risk estimation is also supported

-   [**tidyverse**](https://www.tidyverse.org/), a family of R packages that provide functions to perform common data science tasks including and not limited to data import, data transformation, data wrangling and data visualisation.

# 3 Data Preparation

## 3.1 Downloading the data

The following data sets have been used in this exercise and can be downloaded by clicking the links below:

-   [Thailand Road Accident \[2019-2022\]](https://www.kaggle.com/datasets/thaweewatboy/thailand-road-accident-2019-2022) on Kaggle

-   [Thailand Roads (OpenStreetMap Export)](https://data.humdata.org/dataset/hotosm_tha_roads) on HDX.

-   [Thailand - Subnational Administrative Boundaries](https://data.humdata.org/dataset/cod-ab-tha?) on HDX.

## 3.2 Importing the data in R

Make csv sf geometry must be linestring not multipleline detect driver behavior detect enviromental behavior use to subset point events accident causes accident data point

select different regions from wikipedia of the region do the analysis of the regions like the previous exercise rds folder, make subsets in rds date time field posix incident datetime

::: panel-tabset
## Accident Data
Here we import the accident data and wrangle it so it can be used for our analytics.

```{r}
#| eval: false
rdacc_sf = read_csv("data/rawdata/thai_road_accident_2019_2022.csv") %>%
  filter(!is.na(longitude) & longitude != "",
         !is.na(latitude) & latitude != "") %>%
  st_as_sf(coords = c(
    "longitude", "latitude"),
    crs=4326) %>%
  st_transform(crs = 32647) %>%
  mutate(HourOfDay = hour(incident_datetime))%>%
  mutate(Month_num = month(incident_datetime)) %>%
  mutate(Month_fac = month(incident_datetime,
                       label = TRUE,
                       abbr = TRUE)) %>%
  mutate(dayofweek_num = weekdays(incident_datetime))%>%
mutate(dayofweek_fac = as.numeric(format(incident_datetime, "%u")))
  

```

## Administrative area
We import the following dataset as it contains all sub sections of the administrative area aswell: tha_admbnda_adm3_rtsd_20220121

```{r}
#| eval: false
adm = st_read(dsn = "data/rawdata/", 
                  layer = "tha_admbnda_adm3_rtsd_20220121")
```

And we transform the data to the right CRS

```{r}
#| eval: false
adm = st_transform(adm, 32647)
```

Here we show the dataset
```{r}
#| eval: false
adm
```

## Roads
And the roads.
```{r}
#| eval: false
roads = st_read(dsn = "data/rawdata/", 
                 layer = "hotosm_tha_roads_lines_shp")


```

:::

## MAKING RDS FILES
Now we are creating and reading RDS files for different datasets, specifically for accidents, administrative boundaries, and roads. Also we're performing some data manipulation on the roads dataset, such as filtering and selecting specific columns, and setting the coordinate reference system (CRS) to 4326.

::: panel-tabset
## Accidents

```{r}
#| eval: false
write_rds(rdacc_sf, "data/rds/rdacc_sf.rds")
```

```{r}

rdacc_sf = read_rds("data/rds/rdacc_sf.rds")
```

## Adm

```{r}
#| eval: false
write_rds(adm, "data/rds/adm.rds")
```

```{r}
adm = read_rds("data/rds/adm.rds")
```

## Roads

```{r}
#| eval: false
write_rds(roads, "data/rds/roads.rds")
```

```{r}
#| eval: false
roads = read_rds("data/rds/roads.rds")
```


```{r}
#| eval: false
useroads = c("motorway","primary","secondary")
network = subset(roads, highway %in% useroads)
network <- subset(network, select = -c(name, surface, smoothness, width, lanes, oneway, bridge, layer, source, name_th))
network <- st_set_crs(network, 4326)


```

:::


Taking a "glimpse" at the data

::: panel-tabset
## Accidents

```{r}
#| eval: false
glimpse(rdacc_sf)
```

## Administrative Area

```{r}
glimpse(adm)
```

## Roads

```{r}
#| eval: false
glimpse(network)
```
:::

# Thailand

```{r}
#| eval: false
plot(adm$geometry)
```

our data still has all the provinces of thailand which we dont need, we only need the bankok metropolitan region. This means we will have to exclude some of the data.

First we create a list of the Bankok Metropolan area

```{r}
BMRC = c("Bangkok", "Nakhon Pathom", "Nonthaburi", "Pathum Thani", "Samut Prakan", "Samut Sakhon")
```

then we create a subset of the dataset to focus our data to our use case

```{r}
BMR = subset(adm, ADM1_EN %in% BMRC)
```

We can do the same for the accidents dataset since we don't need the data from outside the region of study

```{r}
BMRACC = subset(rdacc_sf, province_en %in% BMRC)
```

And we write both smaller data sets to our rds folder BMR

```{r}
#| eval: false
write_rds(BMR, "data/rds/BMR.rds")
```
```{r}
read_rds("data/rds/BMR.rds")
```
## fixing roads
We use st union on the administrative boundary to get the outline of the BMR
```{r}
u = st_union(BMR)
u = st_transform(u, 4326)
plot(u)
```

```{r}
#| eval: false
# assume 'roads' is your multilinestring data and 'study_area' is your polygon data
network <- st_intersection(network , u)
```

```{r}
#| eval: false
write_rds(network, "data/rds/network.rds")
```

```{r}
##| eval: false
network = read_rds("data/rds/network.rds")
```

## Roads plotted
```{r}
plot(network$geometry)
```

## Accidents

```{r}
#| eval: false
write_rds(BMRACC, "data/rds/BMRACC.rds")
```

## The Bangkok Metropolitan Region
The Bangkok Metropolitan Region is the urban region surrounding the metropolis of Bangkok. The area covers an area of 7,762 km2 and has a population of 17,400,000.

Different parts of the region
There region consists out of 6 subregions named: Bangkok, Nothaburi, Nakhon Pathom, Pathum Thani, Samut Prakan, and Samut Sakhon

```{r}
plot(BMR["ADM1_EN"])
```

We can also make sub sets of the different parts of the region

```{r}
bk <- BMR %>%
  filter(ADM1_EN == "Bangkok")
nhb <- BMR %>%
  filter(ADM1_EN == "Nonthaburi")
np <- BMR %>%
  filter(ADM1_EN == "Nakhon Pathom")
pt <- BMR %>%
  filter(ADM1_EN == "Pathum Thani")
sp <- BMR %>%
  filter(ADM1_EN == "Samut Prakan")
ss <- BMR %>%
  filter(ADM1_EN == "Samut Sakhon")
```

::: panel-tabset
## Bangkok

```{r}
plot(st_geometry(bk))
```

## Nonthaburi

```{r}
plot(st_geometry(nhb))
```
## Nakhon Pathom
```{r}
plot(st_geometry(np))
```
## Pathum Thani

```{r}
plot(st_geometry(pt))
```

## Samut Prakan

```{r}
plot(st_geometry(sp))
```

## Samut Sakhon

```{r}
plot(st_geometry(ss))
```
:::

Now we look at the different accidents in the BMR

```{r}
tmap_mode('plot')
tm_shape(BMR)+
  tm_polygons() +
tm_shape(BMRACC) +
  tm_dots(size = 0.1, col = "red") 

```
Re-assigning the EPSG code

::: pannel-tabset

## BMR
```{r}
BMR32647 <- st_set_crs(BMR, 32647)
BMR32647 = st_transform(BMR32647, crs = 32647)
```
## ACC
```{r}
BMRACC32647 <- st_set_crs(BMRACC, 32647)
BMRACC32647 = st_transform(BMRACC32647, crs = 32647)
```
## Network
```{r}
network32647 <- st_set_crs(network, 32647)
network32647 = st_transform(network32647, crs = 32647)

```
:::

```{r}
BMR_ACC = st_join(BMR32647, BMRACC32647, by = c("ADM1_EN" = "province_en"))
```

```{r}
tmap_mode("plot")
qtm(BMR_ACC, 
    fill = "number_of_injuries")
```

```{r}
tm_shape(BMR_ACC)+
  tm_polygons("number_of_vehicles_involved")

```

# Spatio-Temporal Point Patterns Analysis

## Overview

### The research questions

The specific questions we would like to answer are:

are the locations of the traffic accidents in Bangkok Metropolitan Region spatial and spatio-temporally independent? if the answer is NO, where and when the observed traffic accidents locations tend to cluster?


### Converting OWIN

Next, as.owin() is used to convert BMR into an owin object.

```{r}
bmr_owin = as.owin(BMR)
bmr_owin
```

Next, class() is used to confirm if the output is indeed an owin object.

```{r}
class(bmr_owin)
```

## Preparing Traffic Accident Data

Above we already have imported the traffic accident data and made it focus on our study area

## Visualising the Traffic Accidents

### Overall Plot

This plot shows all traffic accidents in the region

```{r}
tmap_mode('plot')
tm_shape(BMR)+
  tm_polygons() +
tm_shape(BMRACC) +
  tm_dots()
```

### Visuaising geographic distribution of traffic accidents by month

```{r}
tm_shape(BMR)+
  tm_polygons() +
tm_shape(BMRACC) +
  tm_dots(size = 0.1) +
tm_facets(by="Month_fac", 
            free.coords=FALSE, 
            drop.units = TRUE)
```

## Computing STKDE by Month/Season

In this section, we will compute STKDE by using spattemp.density() of sparr package.

Meteorological Seasons Spring: March-May Summer: June-August Autumn: September-November Winter: December-February

::: panel-tabset
### Extracting traffic accidents by month

The code chunk below is used to remove the unwanted fields from fire_sf sf data.frame. This is because as.ppp() only need the mark field and geometry field from the input sf data.frame.

```{r}
acc_month <- BMRACC %>%
  select(Month_num)
```

### Creating ppp

The code chunk below is used to derive a ppp object called acc_month from acc_month sf data.frame.

```{r}
acc_month_ppp = as.ppp(acc_month)
acc_month_ppp
```

```{r}
summary(acc_month_ppp)
```

```{r}
any(duplicated(acc_month_ppp))
```

```{r}
sum(multiplicity(acc_month_ppp) > 1)
```

JITTERING

```{r}
acc_month_ppp_jit = rjitter(acc_month_ppp,
                            retry =TRUE,
                            nsim=1,
                            drop=TRUE)
```

```{r}
any(duplicated(acc_month_ppp_jit))
```

### Including Owin Object

The code chunk below is used to combine ppp and owin objects into one.

```{r}
acc_month_owin = acc_month_ppp[bmr_owin]
summary(acc_month_owin)
```

As a good practice, plot() is used to plot ff_owin so that we can examine the correctness of the output object.

```{r}
plot(acc_month_owin)
```

### Computing Spatio-temporal KDE

Next, spatttemp.density() of sparr package is used to compute the STKDE.

```{r}
st_kde = spattemp.density(acc_month_owin)
summary(st_kde)
```

### Plotting the spatio-temporal KDE object

In the code chunk below, plot() of R base is used to the KDE 
### Spring

```{r}
tims <- c(3,4,5)
#par(mfcol=c(4,3))
for(i in tims){ 
  plot(st_kde, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at month",i))
}

```
:::
::: panel-tabset
### Summer

```{r}
tims <- c(6,7,8)
#par(mfcol=c(2,3))
for(i in tims){ 
  plot(st_kde, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at month",i))
}

```

### Autumn

```{r}
tims <- c(9,10,11)
#par(mfcol=c(2,3))
for(i in tims){ 
  plot(st_kde, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at month",i))
}

```

### Winter

```{r}
tims <- c(12,1,2)
#par(mfcol=c(2,3))
for(i in tims){ 
  plot(st_kde, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at month",i))
}

```
:::

## Computing STKDE by traffic accidents by day of week

In this section, we will compute the STKDE of traffic accidents by

### Creating ppp object

In the code chunk below, DayofWeek field is included in the output ppp object.

```{r}
acc_wday_ppp <- BMRACC %>% 
  select(dayofweek_fac) %>%
  as.ppp()
```

### Including Owin object

Next, code chunk below is used to combine the ppp object and the owin object.

```{r}
acc_wday_owin <- acc_wday_ppp[bmr_owin]
summary(acc_wday_owin)
```

```{r}
kde_wday <- spattemp.density(
  acc_wday_owin)
summary(kde_wday)
```

```{r}
plot(kde_wday)
```


## Computing STKDE by time of day

In this section, we will compute STKDE by using spattemp.density() of sparr package.


::: panel-tabset
### Extracting traffic accidents by Time of day

The code chunk below is used to remove the unwanted fields from hod_sf sf data.frame. This is because as.ppp() only need the mark field and geometry field from the input sf data.frame.

```{r}
acc_hod <- BMRACC %>%
  select(HourOfDay)
```

### Creating ppp

The code chunk below is used to derive a ppp object called acc_month from acc_month sf data.frame.

```{r}
acc_hod_ppp = as.ppp(acc_hod)
acc_hod_ppp
```

```{r}
summary(acc_hod_ppp)
```

```{r}
any(duplicated(acc_hod_ppp))
```

```{r}
sum(multiplicity(acc_hod_ppp) > 1)
```

JITTERING

```{r}
acc_hod_ppp_jit = rjitter(acc_hod_ppp,
                            retry =TRUE,
                            nsim=1,
                            drop=TRUE)
```

```{r}
any(duplicated(acc_hod_ppp_jit))
```

### Including Owin Object

The code chunk below is used to combine ppp and owin objects into one.

```{r}
acc_hod_owin = acc_hod_ppp[bmr_owin]
summary(acc_hod_owin)
```

As a good practice, plot() is used to plot ff_owin so that we can examine the correctness of the output object.

```{r}
plot(acc_hod_owin)
```

### Computing Spatio-temporal KDE

Next, spatttemp.density() of sparr package is used to compute the STKDE.

```{r}
st_kdehod = spattemp.density(acc_hod_owin)
summary(st_kdehod)
```

### Plotting the spatio-temporal KDE object

In the code chunk below, plot() of R base is used to the KDE

```{r}
tims <- c(0,1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18,19,20,21,22,23)
par(mfcol=c(2,3))
for(i in tims){ 
  plot(st_kdehod, i, 
       override.par=FALSE, 
       fix.range=TRUE, 
       main=paste("KDE at Hour",i))
}

```



Geospatial data wrangling

Converting from sf format into spatstat’s ppp format

```{r}
BMRACC_ppp = as.ppp(BMRACC)
BMRACC_ppp
```

```{r}
plot(BMRACC_ppp)
```

```{r}
summary(BMRACC_ppp)
```

```{r}
any(duplicated(BMRACC_ppp))
```

```{r}
#| eval: false
multiplicity(BMRACC_ppp)
```

```{r}
sum(multiplicity(BMRACC_ppp) > 1)
```

```{r}
bmr_owin = as.owin(BMR)
```

```{r}
plot(bmr_owin)
```

```{r}
summary(bmr_owin)
```

```{r}
BMRACCMAP_ppp = BMRACC_ppp[bmr_owin] 
```

```{r}
summary(BMRACCMAP_ppp)
```

```{r}
plot(BMRACCMAP_ppp)
```

```{r}
bk_owin = as.owin(bk)
nhb_owin = as.owin(nhb)
np_owin = as.owin(np)
pt_owin = as.owin(pt)
sp_owin = as.owin(sp)
ss_owin = as.owin(ss)
```

```{r}
accidents_bk_ppp = BMRACC_ppp[bk_owin] 
accidents_nhb_ppp = BMRACC_ppp[nhb_owin]
accidents_np_ppp = BMRACC_ppp[np_owin] 
accidents_pt_ppp = BMRACC_ppp[pt_owin] 
accidents_sp_ppp = BMRACC_ppp[sp_owin]
accidents_ss_ppp = BMRACC_ppp[ss_owin]
```

```{r}
plot(accidents_bk_ppp)
```

```{r}
plot(accidents_bk_ppp, main="Bangkok" )
plot(accidents_nhb_ppp, main="Nonthaburi")
plot(accidents_np_ppp, main="Nakhon Pathom")  
plot(accidents_pt_ppp, main="Pathum Thani") 
plot(accidents_sp_ppp, main="Samut Prakan")
plot(accidents_ss_ppp, main="Samut Sakhon")
```




## Network constraint Spatial Point Pattern Analysis

Network constrained Spatial Point Patterns Analysis (NetSPAA) is a collection of spatial point patterns analysis methods special developed for analysing spatial point event occurs on or alongside network. The spatial point event can be locations of traffic accident or childcare centre for example. The network, on the other hand can be a road network or river network.

to derive network kernel density estimation (NKDE), and
to perform network G-function and k-function analysis



```{r}
# |eval: false
# loading the road network

tm_shape(network) + 
  tm_lines(col = "black") + 
  tm_shape(BMRACC) + 
  tm_dots(col = "red", size = 0.1)+
  tm_shape(BMR)+
  tm_polygons()
```
First we plot the road network and the accidents
```{r}
plot(st_geometry(network))+
plot(BMRACC,add=T,col='red',pch = 19)
```
Then we need to change the network geometry to line string to enable the nkde to be possible.

```{r}
network <- st_cast(network, "LINESTRING")
```
Now we will create the lixels.
```{r}
lixels <- lixelize_lines(network, 
                         700, 
                         mindist = 375)
```

Continuing we create the variable samples.
```{r}
samples <- lines_center(lixels) 
```
And we make sure the data sets are in the same CRS.
```{r}
samples = st_transform(samples, 32647)
```
```{r}
network = st_transform(network, 32647)
```


Now we render the NKDE
Due to the limited processing power of my laptop this process is not fully possible on my system.
```{r}
densities <- nkde(network, 
                  events = BMRACC,
                  w = rep(1, nrow(BMRACC)),
                  samples = samples,
                  kernel_name = "quartic",
                  bw = 300, 
                  div= "bw", 
                  method = "simple", 
                  digits = 1, 
                  tol = 1,
                  grid_shape = c(1,1), 
                  max_depth = 8,
                  agg = 100,
                  sparse = TRUE,
                  verbose = FALSE)
```

```{r}
samples$density <- densities
lixels$density <- densities
```

rescaling to help the mapping

```{r}

# rescaling to help the mapping
samples$density <- samples$density*10000
lixels$density <- lixels$density*10000
```
Now we plot the NKDE and we see the heatmap on the network roadmap

```{r}
tmap_mode('plot')
tm_shape(BMR)+
  tm_polygons(col = "ADM1_EN", lwd = 0.5, title = "Province")+
tm_shape(filter(lixels, density == 0)) +
  tm_lines(col="grey", lwd = 2, alpha = 0.2)+
tm_shape(lixels)+
  tm_lines(col="density", lwd = 2, title.col = "Per km2", palette = "-inferno")+
  
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
tm_layout(title = "Traffic accidents density map",legend.outside = TRUE)

```
::: panel-tabset
## Per month
```{r}
tmap_mode('plot')
tm_shape(BMR)+
  tm_polygons(col = "ADM1_EN", lwd = 0.5, title = "Province")+
tm_shape(BMRACC) +
  tm_dots(size = 0.01)  +
tm_facets(by="Month_fac", 
            free.coords=FALSE, 
            drop.units = TRUE)+
tm_shape(lixels)+
  tm_lines(col="density", lwd = 2, title.col = "Per km2", palette = "-inferno")+
  tm_shape(filter(lixels, density == 0)) +
  tm_lines(col="grey", lwd = 2, alpha = 0.2)+
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
tm_layout(title = "Traffic accidents density map",legend.outside = TRUE)
```
## Per Hour of Day
```{r}
tmap_mode('plot')
tm_shape(BMR)+
  tm_polygons(col = "ADM1_EN", lwd = 0.5, title = "Province")+
tm_shape(BMRACC) +
  tm_dots(size = 0.01)  +
tm_facets(by="HourOfDay", 
            free.coords=FALSE, 
            drop.units = TRUE)+
tm_shape(lixels)+
  tm_lines(col="density", lwd = 2, title.col = "Per km2", palette = "-inferno")+
  tm_shape(filter(lixels, density == 0)) +
  tm_lines(col="grey", lwd = 2, alpha = 0.2)+
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
tm_layout(title = "Traffic accidents density map",legend.outside = TRUE)
```

## Per Day of Week
```{r}


tmap_mode('plot')
tm_shape(BMR)+
  tm_polygons(col = "ADM1_EN", lwd = 0.5, title = "Province")+
tm_shape(BMRACC) +
  tm_dots(size = 0.01)  +
tm_facets(by="dayofweek_num", 
            free.coords=FALSE, 
            drop.units = TRUE)+
tm_shape(lixels)+
  tm_lines(col="density", lwd = 2, title.col = "Per km2", palette = "-inferno")+
  tm_shape(filter(lixels, density == 0)) +
  tm_lines(col="grey", lwd = 2, alpha = 0.2)+
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
tm_layout(title = "Traffic accidents density map",legend.outside = TRUE)
```
:::
# Network Constrained G- and K-Function Analysis
In this section, we are going to perform complete spatial randomness (CSR) test by using kfunctions() of spNetwork package. The null hypothesis is defined as:

Ho: The observed spatial point events (i.e distribution of accidents) are uniformly distributed over a street network in Bangkok Metropolitan Region.

NOTE: due to the following error i was not able to finish the last section: Error in kfunctions(network, BMRACC, start = 0, end = 1000, agg = 2e+05,  : 
  After snapping the points on the network, some of them share the same location.
         To correct it, please consider setting or increasing the value of the parameter agg.
         They will be merged and their weights added)

```{r}
#| eval: false
kfun_BMRACC <- kfunctions(network, 
                             BMRACC,
                             start = 0, 
                             end = 1000, 
                             agg = 200000,
                             step = 50, 
                             width = 50, 
                             nsim = 50, 
                             resolution = 50,
                             verbose = FALSE, 
                             conf_int = 0.05)

```


```{r}
#| eval: false
kfun_BMRACC$plotk
```


The blue line is the empirical network K-function of the accidents in Bangkok Metropolitan Region. The gray envelop represents the results of the 50 simulations in the interval 2.5% - 97.5%. Because the blue line between the distance of 250m-400m are below the gray area, we can infer that the traffic accidents in bmr area resemble regular pattern at the distance of 250m-400m.

```{r}

```
# Conclusion
We have used different techniques to explore and analyse the different road accidents in the Bangkok metropolitan region. From this we can see that most accidents tend to happen at night as well as in the winter months. 

By using the network constrained methods, we can see that most accidents also occur on the bigger roads connecting Bankok with 
```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```
